\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{pgfplots}
\usepackage{siunitx}
\usepackage{cancel}
\usepackage{enumitem}
\usepackage{txfonts}

\pgfplotsset{compat=1.18}

\usepackage[a4paper, top=1cm, bottom=2cm, left=2cm, right=2cm, includehead, includefoot]{geometry}

\begin{document}

\noindent
Physics 4A - Classical Mechanics \hfill Prof. Roger King

\noindent\rule{\textwidth}{0.4pt}

\begin{center}
    \textbf{\LARGE Homework 2} \\
    \vspace{12pt}
    \large Aaron W. Tarajos \\
    \textit{\today}
\end{center}

\noindent\rule{\textwidth}{0.4pt}

\section*{Problem 1}
In the graph below, is there any time, or time interval, for which the following hold? (a)
$v = 0,\ a = 0$; (b) $v = 0,\ a \ne 0$; (c) $v \ne 0, a = 0$; (d) $v > 0,\ a > 0$; (e) $v > 0,\ a < 0$;
(f) $v < 0,\ a < 0$; (g) $v < 0,\ a > 0$.

\subsection*{Solution}

\section*{Problem 2}
The figure below shows the $v$ versus $t$ graphs for cars A and B. At $t = 0$ both are at $x = 0$.
Estimate: (a) where and when they meet again; and (b) their velocities when they meet.

\subsection*{Solution}
To find when and where the cars meet again after $t = 0$, we must find when $\Delta x$ is equal for both cars. Car A has constant velocity which means that $\Delta x$ is given by;

\[
	\Delta x = v_A t
\]
where $v_A$ is the velocity of Car A. Also, we can see that Car B has constant acceleration which means that $\Delta x$ is given by;
\[
	\Delta x = v_{0B} t + \frac{1}{2} a t^2
\]
where $v_{0B}$ is the initial velocity of car B. We set these equations equal to eachother and re-arrange it to solve for time.
\begin{align*}
	v_A t &= v_{0B} t + \frac{1}{2} a t^2 \\
	v_A t &= \left( v_{0B} + \frac{1}{2} a t \right) t \\
	v_A - v_{0B} &= v_{0B} + \frac{1}{2} a t  - v_{0B} \\
	2 \left( v_A - v_{0B} \right) &= \frac{1}{2} a t \cdot 2 \\
	t &= \frac{2 \left( v_A - v_{0B} \right)}{a} \\
\end{align*}
Looking at the graph we can estimate $a$ for car B to be $3\ \unit{\meter\per\second^2}$, then substituting our variables, we find that the cars meet again at;
\[
	\frac{2 \left( 8.00 - 0.00 \right)}{3.00} = \frac{16}{3}\ \unit{\second}
\]
or approximately 5.33 seconds. Then we use the time in one of our original equation for $\Delta x$ and find that the cars meet again at;

\[
	\left( 8 \right) \cdot \left(\frac{16}{3}\right) = \frac{128}{3}\ \unit{\meter}
\]
which is approximately 42.66 meters. Now, we already know that $v_A$ is constant so it will be 8, however, for car B we must use our acceleration estimate $a$ and $t$ to find the velocity.
\begin{align*}
	v_B &= v_{0b} + at \\
	v_B &= 0 + \left( 3 \right) \cdot \left( \frac{16}{3} \right) \\
	v_B &= 16\ \unit{\meter\per\second}
\end{align*}

\section*{Problem 3}
A Honda Fit is initially with a speed of 115 km/h. Find its acceleration and the time taken
to stop given that: (a) it brakes to a stop in 64.0 m; (b) it crashes head-on into a barrier and
crumples by 1.00 m.

\subsection*{Solution}
Given that we are missing time, we can use the following equation to solve for acceleration in both circumstances and then use the acceleration to solve for time.
\begin{align*}
	v^2 &= 2a \Delta x + v_0^2 \\
	v^2  - v_0^2 &= 2a \Delta x \\
	a &= \frac{v^2  - v_0^2}{2\Delta x}
\end{align*}
We have a unit conversion from km/h to m/s so that we have consistent units for distance and so that the timescale matches the distance units.
\[
	115\ \unit{\kilo\meter\per\hour} \cdot 1000\ \unit{\meter\per\kilo\meter} \cdot \frac{1}{3600}\ \unit{\hour\per\second} = 31.944 \unit{\meter\per\second}
\]
Now we subsitute our values into the equation to find that aceleration when braking to a stop is;
\[
	\frac{0 - 31.944^2}{2\left(64.0\right)} = -8.0\ \unit{\meter\per\second^2}
\]
The time elapsed over that distance then is

\[
	t = \frac{2 \Delta x}{v + v_0}
\]
\[
	\frac{2 (64.0)}{0 + 31.944} = 4.0 \unit{\second}
\]

\section*{Problem 4}
An object moves with constant acceleration. At $t = 2.50$ s, the position of the object is
$x = 2.00$ m and its velocity is $v = 4.50$ m/s. At $t = 7.00$ s, $v = -12.0$ m/s. Find: (a) the
position and the velocity at $t = 0$; (b) the average speed from 2.50 s to 7.00 s, and (c) the
average velocity from 2.50 s to 7.00 s.

\section*{Problem 5}
A ball thrown vertically up from the ground rises to height of 24.0 m. How high would it rise
on the moon if given the same initial speed? The acceleration due to gravity on the moon is one-
sixth that on earth.

\section*{Problem 6}
A ball is thrown up from the top of a building 55.0 m high. It rises to a maximum height
of 20.0 m above the roof. (a) When does it land on the ground below? (b) At what velocity
does it land? (c) When is it 20.0 m below the roof?

\section*{Problem 7}
From the $v$ versus $t$ graph below, plot the following graphs: (a) $a$ versus $t$; (b) $x$ versus $t$.
(c) What is the average acceleration for the first 6.0 s? (d) What is the instantaneous
acceleration at 2.0 s? Assume $x = 0$ at $t = 0$.

\section*{Problem 8}
The length of a train is 44.5 m. Its front is 100. m from a pole. It accelerates from rest at
$0.500\ \unit{\meter\per\second^2}$. (a) How long does it take to go past the pole? (b) At what speeds do its front and
rear pass the pole?

\end{document}
